A Comparison of Graph Processing Systems

### 2020-09-04 documentation/paper/sections/graphFormats.tex
rückmeldung 2, noch nicht alles umgesetzt

\subsection{Important Graph Formats}
Since every frameworks uses different graph input formats, we supply a conversion tool capable of translating from EdgeList to the required formats.
Data Sets retreived from KONECT can be directly read and translated.

The following sections explain the output formats of our conversion tool.
\subsubsection{AdjacencyList}
The AdjacencyList and WeightedAdjacencyList formats are used by Ligra and Polymer. The format was initially specified for the Problem Based Benchmark Suite, an open source repository to compare different parallel programming methodologies in terms of performance and code quality \cite{pbbs}.

The file looks as follows
\begin{equation*}
	n, m, o_1, \ldots, o_n, t_1, \ldots, t_m
\end{equation*}
where commas are \texttt{\textbackslash n}. First, $n$ is the number of vertices and $m$ the number of edges in the graph.

The $o_k$ are the so-called offsets. Each vertex $k$ has an offset $o_k$, that describes an index in the following list of the $t_i$.
The $t_i$ are vertex IDs describing target nodes of a directed edge. 
The index $o_k$ in the list of target nodes is the point where edges outgoing from vertex $k$ begin to be declared. So vertex $k$ has the outgoing edges
\begin{equation*}
	(k, t_{o_k}), (k, t_{o_k+1}),\ldots, (k, t_{o_{k+1}-1}).
\end{equation*}

For the WeightedAdjacencyList format, the weights are appended to the end of the file in an order corresponding to the target nodes.

\subsubsection{EdgeList}
The EdgeList format is the most intuitive and one of the most commonly used in online data set repositories. The KONECT database uses this format and thus it is the input format for our conversion tool.

An edge list is a set of directed eges $(s_1,t_1),(s_2,t_2),\ldots$ where $s_i$ is a vertex ID representing the start vertex and $t_i$ is a vertex ID representing the target vertex.
In the format, there is one edges per line and the vertex IDs $s_i, t_i$ are separated with any whitespace character.

For a WeightedEdgeList, the edge weights are appended to each line, again separated by a whitespace character.

\subsubsection{Binary EdgeList}
The binary EdgeList format is used by Gemini.

For $s_i, t_i$ some vertex IDs and $w_i$ the weight of a directed edge $(s_i,t_i, w_i)$, Gemini requires the following input format
\begin{equation*}
	s_1t_1w_1s_2t_2w_2\ldots
\end{equation*}
where $s_i,t_i$ have \texttt{uint32} data type and the optional weights are \texttt{float32}.
Gemini will derive the number of edges from the file size, so there is no file header or anything similar allowed.

\subsubsection{Giraph's I/O formats}
Giraph is capable of parsing many different input and output formats. All of those are explained in Giraph's JavaDoc\footnote{\url{http://giraph.apache.org/apidocs/index.html}}.
Both edge- and vertex-centric input formats are possible.

One can even define their own input graph representation or output format. For the purposes of this paper, we used an existing format similar to AdjacencyList but represented in a JSON-like manner.

In this format, the vertex IDs are specified as \texttt{long} with \texttt{double} vertex values, \texttt{float} out-edge weights.
Each line in the graph file looks as follows
\begin{equation*}
	[s,v_s,[[t_1, w_{t_1}], [t_2, w_{t_2}]...]]
\end{equation*}
with $s$ being a vertex ID, $v_s$ the vertex value of vertex $s$. The values $t_i$ are vertices for which an edge from $s$ to $t_i$ exists. The directed edge $(s,t_i)$ has weight $w_{t_i}$.

There is no surrounding pair of brackets and no commas separating the lines as it would be expected in a JSON format.

>>>>

\section{Important Graph Formats}
Since every frameworks uses different graph input formats, we supply a conversion tool capable of translating from EdgeList to the required formats.
Data Sets retreived from KONECT can be directly read and translated.

//SNAP

The following sections explain the output formats of our conversion tool.
\subsection{AdjacencyGraph}
The AdjacencyGraph and WeightedAdjacencyGraph formats used by Ligra and Polymer are similar to the more popular \emph{compressed sparse rows} format.
The format was initially specified for the Problem Based Benchmark Suite, an open source repository to compare different parallel programming methodologies in terms of performance and code quality \cite{pbbs}.

The file looks as follows
\begin{equation*}
	x, n, m, o_1, \ldots, o_n, t_1, \ldots, t_m
\end{equation*}
where commas are newlines. The files always start with the name of the format i.e. AdjacencyGraph or WeightedAdjacencyGraph in the first line, here shown as $x$.
Followed by $n$, the number of vertices and $m$ the number of edges in the graph.
The $o_k$ are the so-called offsets. Each vertex $k$ has an offset $o_k$, that describes an index in the following list of the $t_i$.
The $t_i$ are vertex IDs describing target nodes of a directed edge. 
The index $o_k$ in the list of target nodes is the point where edges outgoing from vertex $k$ begin to be declared. So vertex $k$ has the outgoing edges
\begin{equation*}
	(k, t_{o_k}), (k, t_{o_k+1}),\ldots, (k, t_{o_{k+1}-1}).
\end{equation*}

For the WeightedAdjacencyGraph format, the weights are appended to the end of the file in an order corresponding to the target nodes.

\subsection{EdgeList}
The EdgeList format is one of the most commonly used in online data set repositories. The KONECT database uses this format and thus it is the input format for our conversion tool.

An edge list is a set of directed edges $(s_1,t_1),(s_2,t_2),\ldots$ where $s_i$ is a vertex ID representing the start vertex and $t_i$ is a vertex ID representing the target vertex.
In the format, there is one edge per line and the vertex IDs $s_i, t_i$ are separated with any whitespace character.

For a WeightedEdgeList, the edge weights are appended to each line, again separated by any number of whitespace characters.

\subsection{Binary EdgeList}
The binary EdgeList format is used by Gemini.
For $s_i, t_i$ some vertex IDs and $w_i$ the weight of a directed edge $(s_i,t_i, w_i)$, Gemini requires the following input format
\begin{equation*}
	s_1t_1w_1s_2t_2w_2\ldots
\end{equation*}
where $s_i,t_i$ have \texttt{uint32} data type and the optional weights are \texttt{float32}.
Gemini derives the number of edges from the file size, so there is no file header or anything similar allowed.

\subsection{Giraph's I/O formats}
Giraph is capable of parsing many different input and output formats. All of those are explained in Giraph's JavaDoc\footnote{\url{http://giraph.apache.org/apidocs/index.html}}.
Both edge- and vertex-centric input formats are possible.
One can even define their own input graph representation or output format. For the purposes of this paper, we used an existing format similar to AdjacencyList but represented in a JSON-like manner.

In this format, the vertex IDs are specified as \texttt{long} with \texttt{double} vertex values and \texttt{float} out-edge weights.
Each line in the graph file looks as follows
\begin{equation*}
	[s,v_s,[[t_1, w_{t_1}], [t_2, w_{t_2}]...]]
\end{equation*}
with $s$ being a vertex ID, $v_s$ the vertex value of vertex $s$. The values $t_i$ are vertices for which an edge from $s$ to $t_i$ exists. The directed edge $(s,t_i)$ has weight $w_{t_i}$.
There is no surrounding pair of brackets and no commas separating the lines as it would be expected in a JSON format.

### 2020-09-04 documentation/paper/sections/overview/overview.tex
rückmeldung 2, noch nicht alles umgesetzt

\section{Overview}
We start with a short overview describing functionality and characteristics of each framework, followed by their required input data formats.

>>>>

\section{Overview}
In this section we describe the functionality and characteristics of several state-of-the-art graph processing frameworks. 
We start with a short overview describing functionality and characteristics of each framework, followed by their required input data formats.

Further

### 2020-09-04 documentation/paper/sections/preliminiaries.tex
rückmeldung 2, noch nicht alles umgesetzt

\subsection{Graphs and Paths}
An \emph{unweighted graph} is the pair $G=(V,E)$ where the \emph{vertex set} is $V\subseteq\mathbb N$ and the $E$ is the \emph{edge set}.
The edge set describes a number of connections or relations between two vertices. Depending on these relations, a graph can be directed or undirected. For a \emph{directed graph} the edge set becomes 
\begin{equation*}
  E\subseteq\{(x,y)\,|\, x,y\in V, x\neq y\}
\end{equation*}
and in the \emph{undirected} case $E$ is a set of two-sets 
\begin{equation*}
  E\subseteq\{\{x,y\}\,|\, x,y\in V, x\neq y\}.
\end{equation*}

The main difference is thus, that in the case of a directed graph a connection between $s$ and $t$ is not the same as a connection between $t$ and $s$ -- the direction matters.

Independently of the graph being directed or not, a graph can be \emph{weighted}. In this case the edge set is expanded by a numerical value $E_{\text{weighted}}=E_{\text{unweighted}}\times \mathbb R$ further describing the relation.

The size of a graph is often described in the number of edges $|E|$ because this number is typically much larger than the number of vertices $|V|$.

A \emph{Path} from start $s$ to target $t$ is a set of edges $P\subseteq E^n$ with
\begin{equation*}
   P=((x_0,x_1),(x_1,x_2),\ldots, (x_{n-1},x_n))
\end{equation*}
where all $(x_i,x_{i+1})\in E$ and $x_0=s, x_n=t$.

Thus we call a target $t$ \emph{accessible} from $s$ if a Path from $s$ to $t$ exists.

\subsection{Single-Source Shortest-Paths}
Single-Source Shortest-Paths (SSSP) describes the problem of finding a path along some edges of the input graph from a given start node to a target.

Input to the problem is a weighted graph $G=(V,E)$ and a start node $s\in V$. Output is the shortest possible distance from $s$ to each node in $V$. 
The distance is defined as the sum of edge weights $w_i$ on a path from $s$ to the target.
In the case of a unweighted graph, the distance is often described in \emph{hops}, i.e. the number of edges on a path.

The most common implementations are Dijkstra's algorithm or BellmanFord.

\subsection{Breadth-First Search}
Breadth-first search (BFS) is a search problem on a graph. It usually requires an unweighted graph and a start vertex as input. In some cases a target vertex is also given.

The output is usually a set of vertices that are accessible from the start vertex. In the case of a target being given, the output is true if a path from start to target exists or false otherwise.

It is called breadth-first search because the algorithm searches in a path length-based way. First all paths of length 1 i.e. all neighbors of the start node are checked before checking paths of length 2 and so on.

The search algorithm, where the paths of maximum length are checked first is called Depth-First Search.

@@

For example, Google Search uses PageRank to rank web pages in their search engine results.

@@

\paragraph{Delta-PageRank}
?

\subsection{Bulk-Synchronous Parallel Model}
\label{sec:bsp}
The Bulk-Synchronous Parallel (BSP) model is a computation model developed by Leslie Valiant \cite{bsp}. It is commonly used in computation environment with large amounts of synchronous computation, for example Giraph is based on this model.

@@

One of the most famous state-of-the-art graph processing systems is Pregel \cite{pregel}. It is based on the BSP computation model.
Pregel and many open-source versions similar to it were built to process large graphs reliably (offering fault tolerance) on large MapReduce infrastructures.

@@

We show Giraph \cite{Giraph} as an example of a system using this framework.

>>>>

\subsection{Graphs and Paths}
An \emph{unweighted graph} is the pair $G=(V,E)$ where the \emph{vertex set} is $V\subseteq\mathbb N$ and the $E$ is the \emph{edge set}.
The edge set describes a number of connections or relations between two vertices. Depending on these relations, a graph can be directed or undirected. For a \emph{directed graph} the edge set becomes 
\begin{equation*}
  E\subseteq\{(x,y)\,|\, x,y\in V, x\neq y\}
\end{equation*}
and in the \emph{undirected} case $E$ is a set of two-sets 
\begin{equation*}
  E\subseteq\{\{x,y\}\,|\, x,y\in V, x\neq y\}.
\end{equation*}
The main difference is thus, that in the case of a directed graph a connection between $s$ and $t$ is not the same as a connection between $t$ and $s$ -- the direction matters.
The size of a graph is defined as the number of edges $|E|$ // CITATION?.
Independently of the graph being directed or not, a graph can be \emph{weighted}. In this case a function $w:E\rightarrow \mathbb R$ is introduced, that maps an edge to a numerical value, further describing the relation.

A \emph{Path} from starting vertex $s$ to target vertex $t$ is a sequence of vertices
\begin{equation*}
	P=(x_1,x_2,\ldots,x_n)\subseteq V^n
\end{equation*}
with the condition $(x_i,x_{i+1})\in E$ for each $i\in\{1,\ldots,n-1\}$ and $x_1=s, x_n=t$.
Thus we call a target $t$ \emph{reachable} from $s$ if a Path from $s$ to $t$ exists.

\subsection{Single-Source Shortest-Paths}
Single-Source Shortest-Paths (SSSP) describes the problem of finding the shortest path from a starting vertex to every other vertex in the input graph.
Input to the problem is a weighted graph $G=(V,E)$ and a start node $s\in V$. Output is the shortest possible distance from $s$ to each node in $V$. 
The distance is defined as the sum of edge weights $w_i$ on a path from $s$ to the target.
In the case of a unweighted graph, the distance is often described in \emph{hops}, i.e. the number of edges on a path.
The most common implementations are Dijkstra's algorithm or BellmanFord.

\subsection{Breadth-First Search}
Breadth-first search (BFS) is a search problem on a graph. It usually requires an unweighted graph and a start vertex as input.
The output is a set of vertices that are reachable from the start vertex.
In some special cases, a target vertex is also given. In the case of a target being given, the output is true if a path from start to target exists or false otherwise.
It is called breadth-first search because the algorithm searches in a path length-based way. First all paths of length 1 i.e. all neighbors of the start node are checked before checking paths of length 2 and so on.
The search algorithm, where the paths of maximum length are checked first is called Depth-First Search.

@@

@@

For example, Google Search uses PageRank to rank web pages in their search engine results.
\paragraph{Delta-PageRank}
This is a optimization of the traditional PageRank algorithm. The PageRank score of a vertex is only updated if the relative change of the PageRank is larger than some delta. In the next iteration only a subset of PageRanks has to be recalculated.

\subsection{Bulk-Synchronous Parallel Model}
\label{sec:bsp}
The Bulk-Synchronous Parallel (BSP) model is a computation model developed by Leslie Valiant \cite{bsp}. It is commonly used in computation environments with large amounts of synchronous computation.

@@

One of the most famous graph processing systems, Pregel \cite{pregel} is based on the BSP computation model.
Pregel and many open-source versions similar to it were built to process large graphs reliably (offering fault tolerance) on large MapReduce infrastructures \cite{Giraph}.

@@

Giraph \cite{Giraph} is an example of a system using this framework.

### 2020-09-04 documentation/paper/sections/results/bfs.tex
rückmeldung 2, noch nicht alles umgesetzt

\subsection{Breadth-first search}
In these figures, Galois with 96 threads is shown. Again, we show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
\begin{figure*}
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_calcTime.png}
		\caption{Calculation times for BFS on a single node}
		\label{fig:singleNodeBFS_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_execTime.png}
		\caption{Execution times for BFS on a single node}
		\label{fig:singleNodeBFS_exec}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_overheadTimeNormalized.png}
		\caption{Overhead time normalized by the graph size in million edges}
		\label{fig:singleNodeBFS_overheadNormalized}
	\end{subfigure}

@@

\subsubsection{Distributed}
\begin{figure}
	\begin{subfigure}{\columnwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedBFS_calcTime.png}
		\caption{Calculation times for distributed BFS}
		\label{fig:distributedBFS_calc}
	\end{subfigure}
	% \hfil
	\begin{subfigure}{\columnwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedBFS_execTime.png}
		\caption{Execution times for distributed BFS}
		\label{fig:distributedBFS_exec}
	\end{subfigure}
	% \hfil
	% \begin{subfigure}{0.3\textwidth}
	% 	\includegraphics[width=\linewidth]{../../plots/distributedBFS_overheadTimeNormalized.png}
	% 	\caption{Overhead time normalized by the graph size in million edges}
	% 	\label{fig:distributedBFS_overheadNormalized}
	% \end{subfigure}
	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing}
	\label{fig:distributedBFS}
\end{figure}
For both the calculation and the execution times, Breadth-First Search shows similar behaviour as the distributed SSSP test case. All measurements can be seen in \autoref{fig:distributedBFS}.

>>>>

\subsection{Breadth-first search}
\subsubsection{Single-Node}
TODO // Hier fehlt noch Giraph, daher noch keine vollständige Auswertung.

In these figures, Galois with 96 threads is shown. Again, we show the impact of Galois' thread count in \autoref{sec:galois_speedup}.
\begin{figure*}
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_calcTime.png}
		\caption{Calculation times for BFS on a single node}
		\label{fig:singleNodeBFS_calc}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_execTime.png}
		\caption{Execution times for BFS on a single node}
		\label{fig:singleNodeBFS_exec}
	\end{subfigure}
	\hfil
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\linewidth]{../../plots/singleNodeBFS_overheadTimeNormalized.png}
		\caption{Overhead time normalized by the graph size in million edges}
		\label{fig:singleNodeBFS_overheadNormalized}
	\end{subfigure}

@@

\subsubsection{Distributed}
\begin{figure}
	\begin{subfigure}{\columnwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedBFS_calcTime.png}
		\caption{Calculation times for distributed BFS}
		\label{fig:distributedBFS_calc}
	\end{subfigure}
	% \hfil
	\begin{subfigure}{\columnwidth}
		\includegraphics[width=\linewidth]{../../plots/distributedBFS_execTime.png}
		\caption{Execution times for distributed BFS}
		\label{fig:distributedBFS_exec}
	\end{subfigure}
	% \hfil
	% \begin{subfigure}{0.3\textwidth}
	% 	\includegraphics[width=\linewidth]{../../plots/distributedBFS_overheadTimeNormalized.png}
	% 	\caption{Overhead time normalized by the graph size in million edges}
	% 	\label{fig:distributedBFS_overheadNormalized}
	% \end{subfigure}
	\caption{Average times on the distributed cluster, black bars represent one standard deviation in our testing}
	\label{fig:distributedBFS}
\end{figure}
For both the calculation and the execution times, Breadth-First Search shows similar behaviour as the distributed SSSP test case. This is expected since both are graph traversal algorithms starting in one source vertex.
Hence, calculation complexity for each node and communication overhead is similar.
All measurements can be seen in \autoref{fig:distributedBFS}.

### 2020-09-04 documentation/paper/sections/results/galoisSpeedup.tex
rückmeldung 2, noch nicht alles umgesetzt

\subsubsection{Single-source Shortest-path}
\begin{figure}
	\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
	\caption{Calculation time speedup with increasing thread count for Galois Single-source Shortest-paths}
	\label{fig:galoisSpeedupSSSP}
\end{figure}
Starting with SSSP which is the algorithm that really is at an advantage when using many threads in \autoref{fig:galoisSpeedupSSSP}.

For all larger graphs, speedup is in most cases very close to optimal up to about 8 threads. 
Twitter has the best speedup, requiring only 38\%\ the calculation time with 2 threads compared to one, 25\%\ using 4 and 12.9\%\ using 8 threads.
Behaviour on friendster is similar with 52\%\ at 2 threads, 29\%\ at 4 and 16\%\ at 8 threads compared to one.

@@

On all graphs, the speedup never exceeds 6$\times$ even when using 96 thread.
The initial speedup when switching from one to two threads is actually smaller than one, thus decreasing performance, on 4 of 7 graphs. Only flickr, twitter and rMat28 can benefit slightly by a speedup of 2\%, 41\%\ and 5\%\ respectively.

>>>>

\subsubsection{Single-source Shortest-path}
\begin{figure}
	\includegraphics[width=\linewidth]{../../plots/singleNodeSSSPGaloisThreads.png}
	\caption{Calculation time speedup with increasing thread count for Galois Single-source Shortest-paths}
	\label{fig:galoisSpeedupSSSP}
\end{figure}
Starting with SSSP, we see an algorithm that benefits from many available threads in \autoref{fig:galoisSpeedupSSSP}.
For all larger graphs, speedup is in most cases very close to optimal up to about 8 threads. 
Twitter has the best speedup, requiring only 38\%\ the calculation time with 2 threads compared to one, 25\%\ using 4 and 12.9\%\ using 8 threads.
Behaviour on friendster is similar with 52\%\ at 2 threads, 29\%\ at 4 and 16\%\ at 8 threads compared to one.

@@

On all graphs, the speedup never exceeds 6$\times$ even when using 96 threads.
The initial speedup when switching from one to two threads is actually smaller than one, thus decreasing performance, on 4 of 7 graphs. Only flickr, twitter and rMat28 can benefit slightly by a speedup of 2\%, 41\%\ and 5\%\ respectively.

### 2020-09-04 documentation/paper/sections/results/results.tex
rückmeldung 2, noch nicht alles umgesetzt

@@

	The errors rendered the code as-is unable to perform calculations, forcing us to fork the repository and modify the source code. A repository with our changes can be found \href{https://github.com/jasc7636/GeminiGraph}{here}\footnote{\url{https://github.com/jasc7636/GeminiGraph}}.

@@

On a plus side, setup of frameworks like Polymer or Ligra was straight forward and did not require any special treatment.

\clearpage

>>>>

\section{Results}

@@

	The errors rendered the code as-is unable to perform calculations, forcing us to fork the repository and modify the source code. Our changes can be found in one of our repositories\footnote{\url{https://github.com/jasc7636/GeminiGraph}}.

@@

On a plus side, setup of the frameworks Polymer and Ligra was straight forward and did not require any special treatment.

### 2020-09-04 documentation/paper/sections/results/sssp.tex
plots with new Giraph data, some text for dist bfs
rückmeldung 2, noch nicht alles umgesetzt

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTimeNormalizedToGalois.png}
	\caption{Overhead times of each framework normalized by the overhead time of Galois Push}
	\label{fig:distributedSSSP_overhead}
\end{figure}

@@

Both Galois implementations have significantly smaller execution times compared to Gemini or Giraph on all graphs, see \autoref{tbl:ssspexec}.
You can see Gemini being worse by at least a factor of 4 compared to Galois Push on all graphs except flickr.
Giraph's execution times in comparison to this are even worse, taking at least 7$\times$ longer than Galois Push on all graphs.

>>>>

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{../../plots/distributedSSSP_overheadTime.png}
	\caption{Distributed Single-Source Shortest-Paths overhead times of each framework}
	\label{fig:distributedSSSP_overhead}
\end{figure}

@@

Both Galois implementations have significantly smaller execution times compared to Gemini or Giraph on all graphs (cf. \autoref{tbl:ssspexec}).
You can see Gemini being worse by at least a factor of 4 compared to Galois Push on all graphs except flickr.
Giraph's execution times in comparison to this are even worse, taking at least 7$\times$ longer than Galois Push on all graphs.

### 2020-09-04 documentation/paper/sections/testingMethods.tex
rückmeldung 2, noch nicht alles umgesetzt

\section{Testing Methods}
For testing the graph processing systems, we used 5 machines with two AMD EPYC 7401 (24-Cores) and 256 GB of RAM each. One of those machines was only used as part of the distributed cluster, since it only has 128 GB of RAM.
All five machines were running Ubuntu 18.04.2 LTS.

Setup of each framework was performed according to our provided installation guides available in Appendix \ref{app:installationGuides}.
All benchmark cases were initiated by our benchmark script available in our repository. 
All five frameworks are tested on a single server. 
Galois, Gemini and Giraph were benchmarked in on the distributed 5-node cluster as well.
Since Galois supports this parameter, we ran multiple tests comparing Galois' performance with different thread counts on a single machine.

The complete benchmark log files and extracted raw results are available in our repository.

@@

\subsection{Measurements}
For every framework, we measured the \emph{execution time} as the time from start to finish of the console command.

For the \emph{calculation time}, we tried to extract only the time the framework actually executed the algorithm. We came up with the following:
\begin{itemize}
	\item For Galois, we resulted to extracting console log time stamps. Galois outputs \texttt{Reading graph complete.}. Calculation time is the time from this output to the end of execution.

	We know that this is not the most realiable way for measuring the calculation times. 
	Not only due to unavoidable buffering in the console output we expect the measured time to be larger than the actual.
	First, it is not clear that all initialization is in fact complete after reading the graph. Second, we include time in the measurement that is used for cleanup after calculation.

@@

	\item Giraph has built in timers for the iterations (supersteps), the sum of those is the computation time.
\end{itemize}
Furthermore, the \emph{overhead} is the time difference between execution time and calculation time.

Each test case consisting of graph, framework and algorithm was run 10 times, allowing us to smooth slight variations in the measured times.
Later on, we provide the mean values of the individual times as well as the standard deviation where meaningful.

\subsection{Algorithms}
The three problems Breadth-first search (BFS), PageRank (PR) and Single-source shortest-path (SSSP) were used to benchmark framework with every graph.
We always show the results of PageRank with a maximum of five iterations.
For frameworks that support multiple implementations (i.e. PageRank in push and pull modes), we included the alternatives in our testing.

>>>>

\section{Evaluation}
Einleitung

\subsection{Testing Methods}
For testing the graph processing systems, we used 5 machines with two AMD EPYC 7401 (24-Cores) and 256 GB of RAM each. One of those machines was only used as part of the distributed cluster, since it only has 128 GB of RAM.
All five machines were running Ubuntu 18.04.2 LTS.

Setup of each framework was performed according to our provided installation guides available in Appendix \ref{app:installationGuides}.
All benchmark cases were initiated by our benchmark script available in our repository. 
All five frameworks are tested on a single server. 
Galois, Gemini and Giraph were benchmarked in on the distributed 5-node cluster as well.
Since Galois supports this parameter, we ran multiple tests comparing Galois' performance with different thread counts on a single machine.
Unless mentioned otherwise, we always show results of each framework utilizing 96 threads (i.e. the maximum on our machines) for the single-node evaluation.

The complete benchmark log files and extracted raw results are available in our repository\footnote{\url{https://github.com/SerenGTI/Forschungsprojekt} // das ist hier unschön.}.

@@

\subsection{Measurements}
For every framework, we measured the \emph{execution time} as the time from start to finish of the console command.
For the \emph{calculation time}, we tried to extract only the time the framework actually executed the algorithm. 
Furthermore, the \emph{overhead} is the time difference between execution time and calculation time.
For measuring the calculation time, came up with the following:
\begin{itemize}
	\item For Galois, we extract console log time stamps. Galois outputs \enquote{\texttt{Reading graph complete.}}. Calculation time is the time from this output to the end of execution.

	This is not the most realiable way for measuring the calculation times. 
	Not only due to unavoidable buffering in the console output we expect the measured time to be larger than the actual.
	First, it is not clear that all initialization is in fact complete after reading the graph. Second, we include time in the measurement that is used for cleanup after calculation.

	However, this method is the only way of retreiving any measurements without introducing custom modifications to the Galois source code.

@@

	\item Giraph has built in timers for the iterations (supersteps), the sum of those is the computation time.
\end{itemize}
Each evaluation consisting of graph, framework and algorithm was run 10 times, allowing us to smooth slight variations in the measured times.
Later on, we provide the mean values of the individual times as well as the standard deviation where meaningful.

\subsection{Algorithms}
The three problems Breadth-first search (BFS), PageRank (PR) and Single-source shortest-path (SSSP) were used to benchmark framework with every graph.
We always show the results of PageRank with a maximum of five iterations.
For frameworks that support multiple implementations (i.e. PageRank in push and pull modes), we included both in our evaluation.
